---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# textphonemizer

<!-- badges: start -->
![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)
![R version](https://img.shields.io/badge/R-4.0%2B-success)
<!-- badges: end -->

Use textphonemizer to convert natural language text into phonetic spellings using the International Phonetic Alphabet (IPA).  
This package includes tools for:
- Scraping phonetic spellings from Wiktionary
- Converting sentences & text strings to IPA
- Comparing phoneme-level differences (e.g., substitutions, insertions, deletions)
- Building and expanding a custom phonetic dictionary

## Installation

You can install the development version of textphonemizer like so:

``` r
# install from GitHub
# remotes::install_github("saheffley/textphonemizer")
```

## Example

```{r example}
# create dictionary
textphonemizer::generate_dictionary_from_csv("sample_data.csv", text_columns = c("auto_transcript", "manual_transcript"))

# manually add words
textphonemizer::add_to_dictionary("word")
textphonemizer::add_to_dictionary(c("you", "can", "also", "add", "multiple", "words", "at", "once"))

# compare two IPA strings
textphonemizer::compare_phonetic_sequences("he went over the bridge", "he went over the ridge")

# view alignment details
textphonemizer::compare_phonetic_alignment("the cat sat on the mat", "the cat sat in the mat")

```
Whenever you add words to the dictionary, R will print a tibble with only the new words.
