---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(textphonemizer)
```

# textphonemizer

<!-- badges: start -->
![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)
![R version](https://img.shields.io/badge/R-4.0%2B-success)
<!-- badges: end -->

Use textphonemizer to convert natural language text into phonetic spellings using the International Phonetic Alphabet (IPA).  
This package includes tools for:

* Scraping phonetic spellings from Wiktionary
* Converting sentences & text strings to IPA
* Comparing phoneme-level differences (e.g., substitutions, insertions, deletions) based on the Levenshtein distance
* Building and expanding a custom phonetic dictionary

## Installation

You can install the development version of textphonemizer like so:

``` r
# install from GitHub
# remotes::install_github("saheffley/textphonemizer")
```

## Example

```{r reset-dictionary, include = FALSE}
# Wipe the dictionary silently
if (file.exists("phonetic_dictionary.csv")) {
  file.remove("phonetic_dictionary.csv")
}
```

### Adding Words to Local Dictionary
```{r example}
# create dictionary

textphonemizer::generate_dictionary_from_csv("sample_data.csv", text_columns = c("auto_transcript", "manual_transcript"))
# add words

textphonemizer::add_to_dictionary("word")
textphonemizer::add_to_dictionary(c("you", "can", "also", "add", "multiple", "words", "at", "once"))
# manually add custom words (for names, invented words, etc)

textphonemizer::add_manual_ipa("Heffley", "ˈhɛf.li")
```

```{r, include=FALSE}
textphonemizer::add_to_dictionary("she")
```

### Analyzing Strings
```{r more examples}
# compare two IPA strings

textphonemizer::compare_phonetic_sequences("he went over the bridge", "he went over the ridge")
# do more detailed phoneme-level comparisons with error type counts:

textphonemizer::compare_phonetic_alignment("the cat sat on the mat", "the cat sat in the mat")
textphonemizer::compare_phonetic_alignment("she went over the bridge", "he went over the ridge")
```

### Sample Errors
```{r even more examples, warning=TRUE, message=TRUE}
# Some sample errors

textphonemizer::compare_phonetic_alignment("he went on the bridge", "she went on the barge")
textphonemizer::add_to_dictionary("bleepbloop", "airplane")

```
